---
title: "Project Report - MAS8405"
author: "Lloyd Bates"
date: "25/02/2022"
output: pdf_document
---

```{r load, warning=TRUE, include=FALSE}
# Libraries used
library(ProjectTemplate)
library(tidyverse)
library(rjags)
library(coda)
library(ggplot2)
# Please set the working directory to the root folder "Bayesian_Analysis"
setwd("C:/Users/Lloyd Bates/Desktop/Postgraduate Newcastle Data Science MSc/Bayesian Data Analysis/Project/Bayesian_Analysis")
load.project()
```

# Data Description
## Original Data
```{r originalData}
# First few entries of data
head(as.data.frame(Reisby))
```
The data set contains 250 observations of 7 variables: "id","hd","week","lnimi" "lndmi","female","reactive_depression". Here one of the main variables that we will be observing is "lnimi" which represents the log concentration the antidepressant drug Imipramine (IMI) in a patients blood. The question of this report is observe the effectiveness of this drug on a patients depression.

## Data Preprocessing
### 01-A File
Before analysis can be completed it is important to format the data correctly, one detail about the data set is that each row is not a unique person in total there are:  
```{r groups}
# Number of different patients
groups = unique(reisby$id)
length(groups)

# Re-encode id vector
reisby$id = match((reisby$id), groups)
```
This number means that each person was not measured every week. As well the id column has been modified to go from 1:66. Further data transformations have taken place including: converting the week column to a 0/1 representing a placebo week or a week where the drug was administered, also, the Hamilton depression index has been encoded to show the 4 possible levels of depression. Finally, for both a normalized and raw format, the data as been split into test and train data. Note groups are not split as train[200] is 53 and test[201] = 54

```{r newData}
head(reisby)
```

Next the structure of the data frame can be viewed to see if how the variables are stored: 

```{r str}
# View how data is stored
str(reisby)
```

# Exploritory Data Analysis
Firstly to view any relationships between the variables the sactterplotmattrix can be viewed: 

```{r pairs}
# View scatterplot matrix
pairs(reisby)
```

# Regression to Predict Efficacy of Imipramine
Fit a multiple linear regression to the reisby data set and report the posterior mean and a 95% HPD interval for each parameter. To test the model different priors will be used with both a standardized and raw version of the data. We will use the lnimi as the response.

## Jags Code
For the first multiple linear regression the JAGS code is: 

```{r JAGS1}
modelstring="
  model {
    k = 10^3
    b0 ~ dnorm(0, k)
    
    for (j in 1:p) {
      b[j] ~ dnorm(0, k)
    }
    
    tau ~ dgamma(0.001, 0.001)
    sd = pow(tau, -0.5)
    
    for (i in 1:N) {
      y[i] ~ dnorm(mu[i], tau)
      mu[i] = b0 + inprod(b, x[i,])
    }
}"
```

### Prior Selection
For the first model the prior used will be a uninformed, normal distribution as to make b’s to be almost flat.

## R Code
```{r code1}
y = reisby$lndmi
x = reisby[,-4]
x = x[,-1]

jags_data = list(y=y, x=x, N=nrow(x), p=ncol(x))
model = jags.model(textConnection(modelstring), data=jags_data, n.chains=4)
update(model, n.iter=1000)
samples = coda.samples(model, variable.names=c("b0", "sd", "b"),n.iter=1000)
```

## MCMC Diagnostics
Note for this first model only will the diagnostics be shown, it can be assumed that in all following models the same diagnostics will be used.

```{r diag1}
# ACF
mcmc_mat = as.matrix(samples[[1]])
par(mfrow=c(3,3))
for (i in 1:7) {
  acf(mcmc_mat[,i], lag.max=60, main=colnames(mcmc_mat)[i])
}
crosscorr.plot(samples)
```

From the acf plots it is clear that lag 30 is still significant, so a thinning interval of around 32 should be sufficient, now the new JAGS code will look like: 

```{r JAGS2}
jags_data = list(y=y, x=x, N=nrow(x), p=ncol(x))
model = jags.model(textConnection(modelstring), data=jags_data, n.chains=4)
update(model, n.iter=1000)
th = 32
samples = coda.samples(model,
                       variable.names=c("b0", "sd", "b"), thin=th,
                       n.iter=th*1000)

# View how well the chains mixed
gelman.diag(samples, multivariate = FALSE)

# Trace 
par(mfrow=c(3,3))
plot(samples, auto.layout=FALSE, density=FALSE)
```

## Summary
From the mcmc output the following summaries can be found:
```{r DIC1, echo=FALSE}
dic.samples(model, n.iter=th*1000, thin=th)
mcmc_mean = summary(samples)$statistics[,1]
table = as.data.frame(HPDinterval(samples[[1]])[1:5,])
table['mean'] = mcmc_mean[1:5]
table
```

Posterior density plots based on MCMC output: 
```{r density1}
par(mfrow=c(3,3))
for (i in 1:5) {
    d = density(mcmc_mat[,i]) 
    plot(d, main=colnames(mcmc_mat)[i])
}
```

These show that the posterior distributions are approximately normal.